---
title: "Importing Data"
output: 
  html_document:
    df_print: paged
    toc: true
    toc_float: true
---

```{r, message=FALSE, warning=FALSE, include=FALSE}
library(knitr)
opts_chunk$set(comment = NA)
```

Reading CSVs
------------

To read CSVs, as well as any other plain text tabular file formats, we use the package `readr`. The 
function `read_csv` replaces the function `read.csv` from base R and improves upon it in a number 
of ways:

 1. it is much faster
 2. it does not transform headers
 3. it does not convert character columns into factors

Otherwise, it is used in the same way as `read.csv`

```{r}
library(readr)
read_csv("data/TechnicalProvisions.csv")
```

The function guesses the data type of each column by reading the first 1000 rows. If later values 
do not conform to the guess the problems are reported. The types can be specified explicitly using 
the `col_types` argument.

```{r}
read_csv(
  "data/TechnicalProvisions.csv",
  col_types = cols(
    Synd         = col_integer(),
    YoA          = col_integer(),
    GrossPremium = col_double(),
    NetPremium   = col_double()))
```

By default the function will set any `""` or `"NA"` to `NA`, which is R's way of recording missing 
values. __Note:__ Lloyd's has a "NA" risk code so in that case you will need to set `na = ""` in 
the function call.

`read_csv` has a number of other arguments which may be useful, see `help(read_csv)` for more 
details. For more information about the other functions in the `readr` package see its 
[R Documentation](https://www.rdocumentation.org/packages/readr).

Tibbles
-------

You may have noticed `read_csv` does not return a standard `data.frame`, it returns a `tibble`. A 
tibble is very similar to a data.frame but it has a number advantages. By default printing a tibble 
only shows the first 10 rows. This prevent accidently flooding your screen when printing a large 
table.

A data.frame can be converted to a tibble using the `as_tibble` function.

```{r}
library(tibble)

df <- data.frame(Letters = LETTERS, Numbers = 1:26)
as_tibble(df)
```

Tibbles can be created directly using the `tibble` function. Creating tibbles does not convert 
character vectors into factors, it does not convert columns names, and variables can refer to 
previous variables.

```{r}
tibble(
  Letters = LETTERS, 
  Numbers = 1:26, 
  `:)` = paste0(Letters, Numbers))
```

Tibbles can also be created row-wise using `tribble` (transposed tibble) so they are more readable. 
Column names are identified with a "~".

```{r}
letnum <- tribble(
  ~Letters, ~Numbers,
  "A",      1,
  "B",      2,
  "C",      3,
  "D",      4,
  "E",      5)

letnum
```

The package `tibble` includes some useful functions, in particular `add_row` and `add_column` and 
`has_name`:

Add extra rows to a tibble:

```{r}
add_row(letnum, Letters = "F", Numbers = 6)
```

Add extra columns to a tibble (__note__: Cannot refer to or replace an existing column, for that 
use `dplyr::mutate`):

```{r}
add_column(letnum, Random = runif(n = 5))
```

Check whether a table has a specified column name:

```{r}
has_name(letnum, "Letters")
```

For more information about the other functions in the `tibble` package see its 
[R Documentation](https://www.rdocumentation.org/packages/tibble).

Reading Excel Files
-------------------

Excel files can be read into R using the `readxl` package. By default the function `read_excel` 
reads the first sheet in the workbook and returns the contents of the entire sheet:

```{r}
library(readxl)

read_excel("data/TechnicalProvisions.xlsx")
```

Alternatively the sheet can be specified by providing its name:

```{r}
read_excel("data/TechnicalProvisions.xlsx", sheet = "2016 Submission")
```

The column types can also be specified, if necessary:

```{r}
read_excel("data/TechnicalProvisions.xlsx", col_types = c("numeric", "numeric","numeric", "numeric"))
```

You can also obtain a list of all the sheets in a workbook with the funtion `excel_sheets`:

```{r}
excel_sheets("data/TechnicalProvisions.xlsx")
```

The philosphy behind `readxl` is to get the data into R and perform any further cleaning and 
analysis there. The are other packages that give greater control over Excel files. If you wish to 
read and write to Excel try the package [`openxlsx`](https://www.rdocumentation.org/packages/openxlsx).

Importing from SQL Server
-------------------------

Importing data from databases

Other Sources
-------------

### RDS

RDS stands for R Data Serialization. Any R object can be saved directly to a RDS file so it is fast 
to read and write. If you wish to cache large tables and will only be reading them into R this is 
the format to use. The package `readr` contains functions for reading and writing to the format.

```{r}
library(readr)

read_rds("data/TechnicalProvisions.rds")
write_rds(letnum, "data/letnum.rds")
```

### JSON

JSON stands for JavaScript Object Notation. It is a lightweight text-based format commonly used 
in web interfaces It is human-readable and especially useful for non-tabular data. JSON syntax 
has four basic rules:

 - Data is in name/value pairs
 - Data is separated by commas
 - Curly braces hold objects
 - Square brackets hold arrays

Some examples of JSON are:

```{json}
{ "name":"John" }

{ "name":"John", "age":31, "city":"New York" }

{ "employees": [
  { "firstName": "John", "lastName": "Doe" },
  { "firstName": "Anna", "lastName": "Smith" },
  { "firstName": "Peter", "lastName": "Jones" }
] }
```

To read JSON files we use the `jsonlite` package, which returns the data in the form of a list.

```{r}
library(jsonlite)

# TODO: Replace with a Lloyd's oriented example
read_json("data/employees.json")

```

### XML

XML stands for eXensible Markup Language. It is also a text-based format for storing non-tabular 
data, which is more verbose than JSON. It is a very popular way to store data for applications.
HTML, used for encoding web pages is a special form of XML.

```{xml}
<name>John</name>

<name>John</name> <age>31</age> <city>New York</city>

<employees>
  <employee>
    <firstName>John</firstName> <lastName>Doe</lastName>
  </employee>
  <employee>
    <firstName>Anna</firstName> <lastName>Smith</lastName>
  </employee>
  <employee>
    <firstName>Peter</firstName> <lastName>Jones</lastName>
  </employee>
</employees>
```

To read XML files we use the `xml2` package, which returns an XML document. To use the data in R 
the simplest method is to convert it to a list using `as_list`.

```{r}
library(xml2)

xml_doc <- read_xml("data/employees.xml")
xml_doc
```

```{r}
as_list(xml_doc)
```

### YAML

YAML stands for YAML Ain't Markup Language. Another popular file format, particularly for 
configuration files. It is very similar to JSON but it is even more lightweight. It uses new lines 
to separate data; indentation to show heirarchy and "-" to indicate array elements. For example:

```{yaml}
name : John
age  : 31
city : New York

employees :
  - firstName : John
    lastName  : Doe
  - firstName : Anna
    lastName  : Smith
  - firstName : Peter
    lastName  : Jones
```

There is currently a "tidyverse" version of an R package but the package `yaml` does the job with 
its function `yaml.load_file`

```{r}
library(yaml)

yaml.load_file("data/employees.yaml")
```
