---
title: "Importing Data"
author: "Chad Goymer"
output: html_document
---

Reading CSVs
------------

```{r}
library(readr)

# read_csv("D:/Temp/NetPremiums_31225.csv")
```

The function guesses the data type of each column by reading the first 1000 rows. If later values 
do not conform to the guess the problems are reported. The types can be specified explicitly using 
the `col_types` argument.

```{r}
library(readr)

# read_csv(
#   "D:/Temp/NetPremiums_31225.csv", 
#   col_types = cols(
#     Synd = col_integer(),
#     YoA = col_integer(),
#     RG = col_character(),
#     Ccy = col_character(),
#     Value = col_double()))

```

By default the function will set any missing values `""` or `"NA"` to `NA`, which is R's way of 
recording missing values. __Note:__ Lloyd's has a "NA" risk code so in that case you will need 
to set `na = ""` in the function call.

`read_csv` has a number of other arguments which may be useful, see `help(read_csv)` for more 
details.

Tibbles
-------

You may have noticed `read_csv` does not return a standard `data.frame`, it returns a `tibble`. A 
tibble is very similar to a data.frame but it has a number advantages. By default printing a tibble 
only shows the first 10 rows. This prevent accidently flooding your screen when printing a large 
table.

A data.frame can be converted to a tibble using the `as_tibble` function.

```{r}
library(tibble)

df <- data.frame(Letters = LETTERS, Numbers = 1:26)
as_tibble(df)
```

Tibbles can be created directly using the `tibble` function. Creating tibbles does not convert 
character vectors into factors, it does not convert columns names, and variables can refer to 
previous variables.

```{r}
library(tibble)

tibble(
  Letters = LETTERS, 
  Numbers = 1:26, 
  `:)` = paste0(Letters, Numbers))
```

Tibbles can also be created row-wise using `tribble` (transposed tibble) so they are more readable. 
Column names are identified with a "~".

```{r}
library(tibble)

tribble(
  ~Letters, ~Numbers,
  "A",      1,
  "B",      2,
  "C",      3,
  "D",      4,
  "E",      5)
```

The package `tibble` includes some useful functions, in particular `add_row` and `add_column`

```{r}
tribble(
  ~Letters, ~Numbers,
  "A",      1,
  "B",      2,
  "C",      3,
  "D",      4,
  "E",      5)
```

Reading Excel Files
-------------------

Importing from SQL Server
-------------------------

Other Sources
-------------

RDS

JSON

XML
